SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-20 paper/bio_MTedit.tex
intro-bio-start

%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}
Since a VTS are a transport graph, it is but natural to formally model VTS as graphs (as used in computer science) with their nodes denoting compartments and labeled edges denoting transport vesicles with labels denoting the set of molecules being transported.  The budding and fusion regulations can be represented as matching tables over sets of molecules, i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met.  Given such a graph model of a VTS, then properties, such as $k$-connectedness or steady state condition, can be formally defined as constraints, i.e., first-order logical formulas, over graphs and function types.  Note that these formulas define among other things constraints over paths between nodes in the graph model of VTS. 
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.

@@

Similarly, one can define fusion and budding rules as constraints over the boolean functions modeling the regulations.  Given definitions of fusion and budding rules and the steady state conditions, whether a VTS meets maximum connectivity requirement, i.e. the LGC property, can be verified by checking if the formula show in Listing~1.2 is valid.  Here we have defined the property by checking over existence of any fusion/budding rules.  We can eliminate the existential quantifier if we are interested in checking the property for a particular fusion/budding rule.

\subsection{Converting the Graph Problem into Boolean SAT problem}
\label{subsec:satproblem}
To convert the graph problem described in sec~\ref{subsec:graphmodeling}, into a boolean SAT problem, we need to define a scheme to represent graphs and boolean functions using a suitable set of propositional variables.  In our earlier work, we modeled the graph problem in C using arrays to model graphs and boolean function.  We then used the CBMC model checker to convert the graph problem into a SAT problem.  One of the challenges we had in our earlier work is dealing with quantifiers.  Note that the connectedness property defined in Listing~1.2 has quantifier alternation.  Even if we were to eliminate the existential quantifier by instantiating the problem for a fixed set of fusion rules, we would have embedded quantifiers in the antecedent of implications.   CBMC supports only a quantifier-free logic n its assertion language.  In our earlier work we used a combination of explicit enumeration at the C-level and clever use f non-determinism to eliminate alternation of quantifiers.  This enumeration was one source of bottleneck for scaling in our earlier work.  In the current work we eliminate this bottleneck by modeling the problem directly as a SAT instance using uninterpreted functions and recursive relations.  The details of te hnew encding will described in subsequent sections.

>>>>

%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}
Since a VTS are a transport graph, it is but natural to formally model VTS as graphs (as used in computer science) with their nodes denoting compartments and labeled edges denoting transport vesicles with labels denoting the set of molecules being transported.  The budding and fusion regulations can be represented as matching tables over sets of molecules.
% i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met. 
Given such a graph model of a VTS, then properties, such as $k$-connectedness or steady state condition, can be formally defined as constraints, 
%i.e., first-order logical formulas, 
over graphs and function types.  Note that these formulas define among other things constraints over paths between nodes in the graph model of VTS. 
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.

@@

Similarly, one can define fusion and budding rules as constraints over the boolean functions modeling the regulations.  
%Given definitions of fusion and budding rules and the steady state conditions, whether a VTS meets maximum connectivity requirement, i.e. the LGC property, can be verified by checking if the formula show in Listing~1.2 is valid.  Here we have defined the property by checking over existence of any fusion/budding rules.  We can eliminate the existential quantifier if we are interested in checking the property for a particular fusion/budding rule.
%
%\subsection{Converting the Graph Problem into Boolean SAT problem}
%\label{subsec:satproblem}
%To convert the graph problem described in sec~\ref{subsec:graphmodeling}, into a boolean SAT problem, we need to define a scheme to represent graphs and boolean functions using a suitable set of propositional variables.  In our earlier work, we modeled the graph problem in C using arrays to model graphs and boolean function.  We then used the CBMC model checker to convert the graph problem into a SAT problem.  One of the challenges we had in our earlier work is dealing with quantifiers.  Note that the connectedness property defined in Listing~1.2 has quantifier alternation.  Even if we were to eliminate the existential quantifier by instantiating the problem for a fixed set of fusion rules, we would have embedded quantifiers in the antecedent of implications.   CBMC supports only a quantifier-free logic n its assertion language.  In our earlier work we used a combination of explicit enumeration at the C-level and clever use f non-determinism to eliminate alternation of quantifiers.  This enumeration was one source of bottleneck for scaling in our earlier work.  In the current work we eliminate this bottleneck by modeling the problem directly as a SAT instance using uninterpreted functions and recursive relations.  The details of te hnew encding will described in subsequent sections.
